Implicit assumptions and their impact on photometric redshift PDF performance in the context of LSST

### 2018-11-27 paper/tex/introduction.tex
more edits to the introduction

There are numerous techniques for deriving photo-$z$ PDFs from photometry, yet no one method has yet been established as clearly superior.
Quantitative comparisons of photo-$z$ methods have been made before.
The Photo-$z$ Accuracy And Testing \citep[PHAT,][]{Hildebrandt:10} effort focused on point estimates derived from many photometric bands.  \citet{Rau:2015} introduced a new method for improving redshift PDFs using an ordinal classification algorithm.
DES compared several codes for point estimates and a subset with $p(z)$ information \citep{Sanchez:14}.  A follup up paper examined summary statistics of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}.  
This paper is distinguished by its focus on metrics of photo-$z$ interim posteriors themselves and consideration of both classic and state-of-the-art photo-$z$ algorithms, comparing the performance of several of the most widely employed codes as well as some that have been developed only recently on the basis of metrics appropriate for a probabilistic data product.
The results presented in this work are
%part of the Data Challenge 1 (DC1)
a major focus of the Photometric Redshift working group of the LSST-DESC.
This work is laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}} as one of the critical activities to be completed in preparation for dark energy science analysis on the first year LSST data.
%This is the first of multiple papers by the working group, which will grow in sophistication.
In this initial paper we focus on evaluating the performance of photometric redshift codes and PDF-based performance metrics in the presence of representative training sets.
Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.  We note that these initial tests are a necessary requirement for photo-$z$ codes that will be used in cosmological analyses; however, meeting these requirements is only the first stage in the process, and can be thought of as an initial test under near perfect conditions before further complexities are added in future analyses.

Before moving forward, we must address how the best methods may be unique to the performance metrics and science cases considered and what distinguishes photo-$z$ PDFs of different methods from one another.
Though photo-$z$ PDFs are often written simply as $p(z)$, the PDF itself must be an interim posterior distribution $p(z | d, I)$, the probability of redshift conditioned on photometric data $d$ that has actually been observed and the prior information $I$ that guides how a redshift is extracted from the photometry.
If we run multiple photo-$z$ codes on a single dataset, the photo-$z$  posteriors will not be identical because each code employs assumptions that constitute an interim prior; these assumptions form the premise for photo-$z$ estimation as a whole and are the only way to introduce differences in estimates of what would otherwise be a shared photo-$z$ posterior $p(z | d)$ regardless of the code used to obtain it.
Though explicit knowledge of the interim prior is necessary to use photo-$z$ posteriors self-consistently in physical inference, the interim prior of a particular methodology is often implicit and not necessarily shared among all galaxies in the catalogue.

This paper therefore aims (1) to constrain the impact of the interim prior $I$ by separating it into a component $I_{H}$ representing the method itself and a component $I_{D}$ representing physical information, such as a training set or SED template library and (2) to present a procedure for evaluating the performance of photo-$z$ codes in generic tests that may include many more systematics in the interim prior $I$.
In order to isolate the effects encapsulated by $I_{H}$ of variation between codes from issues with the training set or template library encapsulated by $I_{D}$, we use an identical set of simulated galaxies for every code and construct a template library and training sample that are {\it complete and representative} and shared among all codes; that is, our training sample for machine learning codes is drawn from the same underlying galaxy population as our test set, with no additional selections, and the SED library used for template-based codes is the same as the one used to generate the photometric data.
We explore a number of performance metrics in this paper, in the controlled environment of the representative training data in order to establish a method for comparing photo-$z$ PDFs derived by different methods.
%not to make a conclusion regarding the superiority or even relative favorability of each code but to establish a method for comparing photo-$z$ PDFs derived by different methods.
These test conditions set the stage for addressing in a future paper the crucial issue of incomplete and non-representative prior information.

>>>>

There are numerous techniques for deriving photo-$z$ PDFs from photometry, yet no one method has yet been established as clearly superior.
Quantitative comparisons of photo-$z$ methods have been made before.
The Photo-$z$ Accuracy And Testing \citep[PHAT,][]{Hildebrandt:10} effort focused on point estimates derived from many photometric bands.  \citet{Rau:2015} introduced a new method for improving redshift PDFs using an ordinal classification algorithm.
DES compared several codes for point estimates and a subset with $p(z)$ information \citep{Sanchez:14}.  A follup up paper examined summary statistics of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}.

This paper is distinguished by its focus on metrics of photo-$z$ interim posteriors themselves and consideration of both classic and state-of-the-art photo-$z$ algorithms, comparing the performance of several of the most widely employed codes as well as some that have been developed only recently on the basis of metrics appropriate for a probabilistic data product.
The results presented here are
%part of the Data Challenge 1 (DC1)
a major focus of the Photometric Redshift working group of the LSST-DESC.
This work is laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}} as one of the critical activities to be completed in preparation for dark energy science analysis on the first year LSST data.
%This is the first of multiple papers by the working group, which will grow in sophistication.
In this initial paper we focus on evaluating the performance of photometric redshift codes and PDF-based performance metrics in the presence of complete and representative training sets.  Specific implementation choices in each code will influence the resultant posterior distributions, for example choice of prior parameterization in template-based codes, the bandwidth size chosen for machine learning based codes, or even the output format chosen for storing the PDF.  We have attempted to minimize the impact of many of these factors when comparing codes, for example by using the same template set for all template-based codes, and using a training set that is drawn from the same underlying population as the test sample, to create a controlled environment in which to compare the photo-$z$ PDFs derived from each method.  We explore a number of performance metrics in this paper that test whether the posterior estimates are actual PDFs.  
Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.  Identifying and fixing failure modes within codes may aid us in reaching the stringent photo-$z$ performance goals set out for LSST.  We note that these initial tests are a necessary requirement for photo-$z$ codes that will be used in cosmological analyses; however, meeting these requirements is only the first stage in the process, and can be thought of as an initial test under near perfect conditions to test for problems before further complexities are added in future analyses.

%Before moving forward, we must address how the best methods may be unique to the performance metrics and science cases considered and what distinguishes photo-$z$ PDFs of different methods from one another.  Though photo-$z$ PDFs are often written simply as $p(z)$, the PDF itself is an interim posterior distribution $p(z | d, I)$, the probability of redshift conditioned on photometric data $d$ that has been observed and the prior information $I$ that guides how a redshift is extracted from the photometry.  If we run multiple photo-$z$ codes on a single dataset, the photo-$z$ posteriors will not be identical because each code employs assumptions that constitute an interim prior (for example, the specific template set used in a template-based code, or the bandwidth/number of neighbors in a machine learning based code).  %These assumptions form the premise for photo-$z$ estimation as a whole and are the only way to introduce differences in estimates of what would otherwise be a shared photo-$z$ posterior $p(z | d)$ regardless of the code used to obtain it.

%Though explicit knowledge of the interim prior is necessary to use photo-$z$ posteriors self-consistently in physical inference, the interim prior of a particular methodology is often implicit and not necessarily shared among all galaxies in the catalogue.

%This paper aims to constrain the impact of these differing interim priors and isolating the effects due to the differing code methodologies by using a homogeneous

%This paper therefore aims (1) to constrain the impact of the interim prior $I$ by separating it into a component $I_{H}$ representing the method itself and a component $I_{D}$ representing physical information, such as a training set or SED template library and (2) to present a procedure for evaluating the performance of photo-$z$ codes in generic tests that may include many more systematics in the interim prior $I$.
%In order to isolate the effects encapsulated by $I_{H}$ of variation between codes from issues with the training set or template library encapsulated by $I_{D}$, we use an identical set of simulated galaxies for every code and construct a template library and training sample that are {\it complete and representative} and shared among all codes; that is, our training sample for machine learning codes is drawn from the same underlying galaxy population as our test set, with no additional selections, and the SED library used for template-based codes is the same as the one used to generate the photometric data.
%We explore a number of performance metrics in this paper, in the controlled environment of the representative training data in order to establish a method for comparing photo-$z$ PDFs derived by different methods.
%not to make a conclusion regarding the superiority or even relative favorability of each code but to establish a method for comparing photo-$z$ PDFs derived by different methods.
%These test conditions set the stage for addressing in a future paper the crucial issue of incomplete and non-representative prior information.
